\documentclass[12pt]{article}

\usepackage[margin=1.1in]{geometry}
\input{../../syllabi/preamble}

%%edit these!!
\newcommand{\coursedept}{Math}
\newcommand{\coursenumber}{369}
\newcommand{\coursenumbercrosslisted}{/ 690.03~}
\newcommand{\semester}{Fall}
\newcommand{\numcredits}{3}
\newcommand{\lectimeandloc}{Tues and Thurs 5 -- 6:15PM / on zoom}
\newcommand{\requiredlabtimeandloc}{}
\newcommand{\tataofficehourtimeandloc}{TA / TA Office Hours / Loc & Kennly Weerasinghe / see \coursewebpagelink}
\newcommand{\numtheoryhws}{6--9}
\newcommand{\lastdatetimetohandinhomeworks}{May 18 at noon}
\newcommand{\midtermonedateandlocation}{Thursday, October 7 on zoom during class time}
\newcommand{\midtermtwodateandlocation}{Thursday, November 11 on zoom during class time}
\newcommand{\finaldateandlocation}{TBD but on zoom}
\newcommand{\sixhundredsection}{690.x }

%%%unchanged
\newcommand{\coursewebpageurl}{https://github.com/kapelner/QC_\coursedept_\coursenumber_\semester_\the\year}
\newcommand{\coursewebpagelink}{\href{\coursewebpageurl}{course homepage}}
\newcommand{\slackurl}{https://QC\coursedept\coursenumber\semester\the\year.slack.com/}
\newcommand{\slacklink}{\href{\slackurl}{slack}}

\input{../../syllabi/_header}



\section*{Course Overview}

MATH 369 is an introduction to the intermediate concepts of mathematical statistics, statistical inference and theory of experimentation. Main topics are listed below:


\begin{itemize}
\itemsep -0.0em 
\item Estimation
\begin{itemize}
\item Data generating processes, parameters, estimators, estimates, samples, population 
\item Point parameter estimators and estimates: loss, risk, bias-variance decomposition
\item Likelihood, log-likelihood, score function
\item Un/biased estimators and asymptotically unbiased estimators, consistent estimator
\item Method of Moments (MM), Maximum likelihood estimator (MLE)
\item Cramer-Rao Lower Bound, Fisher information, UMVUEs
\item Asymptotical normality and efficiency of MLE's, uni/multivariate delta methods
\item The Empirical CDF estimator
\item The non-parametric bootstrap
\end{itemize}
\item Testing
\begin{itemize}
\item Neyman-Pearson paradigm of Hypothesis testing
\item Left-sided/tailed, right-sided/tailed, two-sided/tailed tests
\item One proportion z-test, One sample z-test, one sample t-test, two sample z-test (un/equal variances), two sample t-test  (un/equal variances), two-proportion z-test.
\item Size ($\alpha$), Fisher's $p$-value and \qu{statistical significance}
\item Statistical power and the power function
\item Effect sizes and \qu{practical / clinical significance}
\item uniformly most powerful tests?, equivalence testing?
\item Wald, score and the generalized likelihood ratio testing procedures
\item Chi-squared tests for goodness of fit and independence
\item Kolmogorov-Smirnov's goodness of fit test and distribution equivalence test
\item Permutation Tests
\item Multiple testing comparison problem: familywise error control via the Bonferroni, Sidak and Simes procedures, false discovery rate control via the Benjamini-Hochberg procedure
\item The non-parametric bootstrap
\end{itemize}
\item Experimentation and Causality
\begin{itemize}
\item Casual inference vs statistical inference
\item Neyman-Rubin model, the counterfactual
\item Observational study, confounding variable, biased causal estimation
\item Experimentation, completely randomized experiment, unbiased causal estimation
\end{itemize}
\item Model Selection
\begin{itemize}
\item Naive model selection bias based on number of parameters (overfitting)
\item Model section criterions: AIC, AICC
\item Akaike weights / scores
\end{itemize}
\end{itemize}



\textbf{This is more of a typical mathematics theory course than the rest of the data science series but it is still not like the other math courses you're used to; it will be more philosophical and very high in conceptual density while the mathematics itself is kept deliberately simple.} But we will still attempt to keep our eye on developing ideas and concepts for helping to make decisions in the real world. Thus we may make limited use of computation using the \texttt{R} statistical language.

\subsection*{Prerequisites}

MATH 241 (basic probability), 201 (multivariable calculus) and 231 (linear algebra) or equivalents. I expect a 241 class that covers more or less what I cover in 241. See the course homepage for links under \qu{prerequisite review}. The multivariable calculus and linear algebra we will use I will try to review in class.

\section*{Course Materials}

\paragraph{Textbook:} I will be referencing Larry Wasserman's \emph{All of Statistics: A concise course in statistical inference} which can be purchased on \href{https://www.amazon.com/dp/0387402721}{Amazon} and Casella and Berger's \emph{Statistical Inference} which can be purchased on \href{https://www.amazon.com/dp/8131503941}{Amazon}. There is no excuse not to have these books. They are \textit{required}. However, I will not ususally be teaching \qu{from the book} --- most of the material in the class comes from the lecture notes. The textbooks are a way to get ``another take'' on the material and they will only cover about only half of the material done in class. For the other half, you will have to make use of other resources. I also recommend Rice's \emph{Mathematical Statistics and Data Analysis}, 3rd edition which can be purchased on \href{https://www.amazon.com/dp/0534399428}{Amazon} as well but I will not reference it during class.

\paragraph{Computer Software:} During lectures, there will be demoes using \texttt{R} which is a free, open source statistical programming language and console. You can download it from: \url{http://cran.mirrors.hoobly.com/}. I do not expect you to do \textit{any} programming. If \texttt{R} will be part of the homework, I will give you the code to run and expect you to interpret the results based on concepts explained during the course.

\paragraph{Calculator:} You can use a TI-84, 85, 89 or any calculator which you wish. I strongly suggest you use \href{http://www.wolframalpha.com/}{Wolfram Alpha} and its smartphone app.

\input{../../syllabi/_the650section}

\input{../../syllabi/_announcements_on_slack}

\input{../../syllabi/_use_of_slack}

\input{../../syllabi/_standard_class_meetings}

\input{../../syllabi/_jewish_holiday_reschedule}

\input{../../syllabi/_zoom_policies}

\input{_lecture_schedule}

%\input{../../syllabi/_lecture_upload}

\section*{Homework}

\input{../../syllabi/_theory_hws_text}

\input{../../syllabi/_theory_hws_submission_text}
\input{../../syllabi/_philosophy_hws}

\input{../../syllabi/_time_spent_hws}

\input{../../syllabi/_late_hw_policy}

\input{../../syllabi/_latex_hw_bonus_policy}

\input{../../syllabi/_hw_ec_policy}

\section*{Examinations}

\input{../../syllabi/_examination_text}

\input{../../syllabi/_standard_exam_schedule}

\subsection*{Exam Policies and Materials}

\input{../../syllabi/_examination_policies}

\input{../../syllabi/_zoom_examination_policies}

\input{../../syllabi/_standard_cheat_sheet_policy}


\input{../../syllabi/_cheating_on_exams_and_missing_exams}
\input{../../syllabi/_special_services}

\input{../../syllabi/_class_participation}

%\input{../../syllabi/_zoom_attendance}

\input{../../syllabi/_standard_grading_and_grading_policy}

\input{../../syllabi/_advanced_course_grade_distribution}

\input{../../syllabi/_grade_checking_on_gradesly}

\input{../../syllabi/_auditing_policy}

\end{document}